---
title: "Covid-19-WHO-Data-Exploration -for_INDIA_&_UAE"
author: "sruthy kannath"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a Project done with R markdown (Rstudio) to unserstand the covid 19 data for various countries .I am mainly focusing with my native and residing country (i.e, India and Uae).Here I am using latest covid 19 data from WHO by using Special R library named covid19.analytics.
```{r }
library(covid19.analytics)
```

## Data Extraction using method specified by the libraray

```{r}
covid_data <- covid19.data(case = 'aggregated')
confirmed_case <- covid19.data(case='ts-confirmed')
total_cases <- covid19.data(case='ts-ALL')
```

## Summary of cases with countries

```{r}
report.summary(Nentries = 10,graphical.output = T)
```

## Total for Location INDIA and UAE
```{r}
tots.per.location(confirmed_case,geo.loc='india')
tots.per.location(confirmed_case,geo.loc='UNITED ARAB EMIRATES')


```

## Growth rate of India and UAE
```{r}
growth.rate(confirmed_case,geo.loc='india')
growth.rate(confirmed_case,geo.loc='UNITED ARAB EMIRATES')


```
## plot total cases globaly
```{r}
totals.plt(total_cases)
totals.plt(total_cases,'India')
totals.plt(total_cases,'UNITED ARAB EMIRATES')

```

## ploting total cases in world map

```{r}
live.map(total_cases)
```

## SIR Model
```{r}
growth.rate(confirmed_case,geo.loc='india')
growth.rate(confirmed_case,geo.loc='UNITED ARAB EMIRATES')


```

